Getting start

### 2023-05-02 inst/gitbook/01_getting_started.Rmd
Update 01_getting_started.Rmd

@@

```{r}
## Plotting the results (first trait)
plot(sim_tree, trait = 1)
```

@@

```{r}
## TODO: make this automatically do this:
# time_series <- chrono.subsets(data = dispRitreats(sim_data)$data, tree = dispRitreats(sim_data)$tree, method = "continuous", model = "proximity", time = 13)

## Converting the data to be passed to dispRity functions
dispRity_data <- dispRitreats(sim_data)

## Creating the time series
time_series <- chrono.subsets(data = dispRity_data$data, tree = dispRity_data$tree, time = seq(from = 0.9, to = 0, by = -0.1), method = "continuous", model = "proximity")

simulated_disparity <- dispRity(time_series, metric = c(sum, variances), verbose = TRUE)

```

>>>>

> For more info about `chrono.subsets` and the `dispRity` package, you can have a look at the [`dispRity` manual](http://tguillerme.github.io/dispRity.html).

@@

```{r}
## Plotting the results (first trait)
plot(sim_data, trait = 1)
```

@@

```{r}
## Converting the data to be passed to dispRity functions
dispRity_data <- dispRitreats(sim_data)

## Creating the time series (by default dispRitreats scales the trees to be 1 unit of times old)
time_series <- chrono.subsets(data = dispRity_data$data,
                              tree = dispRity_data$tree,
                              time = seq(from = 0.9, to = 0, by = -0.1),
                              method = "continuous", model = "proximity")

## Calculating the disparity from the simulated data
simulated_disparity <- dispRity(time_series, metric = c(sum, variances),
                                verbose = TRUE)
par(mfrow = c(1,2))
plot(simulated_disparity, main = "simulated disparity")
plot(change_in_size, main = "observed disparity")
```

We can see that both simulated and observed disparity curves are different.
From there one could already reasonably propose that the observe disparity does not entirely follows our parametrised scenario (constant speciation and extinction and BM trait evolution).

However, we can complexify our scenario to introduce the effect of mass extinction event.
We can simulate two alternative scenarios:
 * one where we create a random extinction event after having simulated half of the species (25) where we remove 75% of species;
 * and a second one where we remove all the species with negative values.
This allows us to simulate two contrasted scenarios to see whether using our observed dataset it would be possible to detect changes in disparity due to mass extinctions.

To do so, we need to create two mass extinction `"event"` objects:

```{r}
## Creating a random mass extinction
random_extinction <- make.events(target = "taxa",
                                 condition = taxa.condition(25, condition = `>=`),
                                 modification = random.extinction(0.75))
## Creating an extrinction that removes species with negative trait values
negative_extinction <- make.events(target = "taxa",
                                   condition = taxa.condition(25, condition = `>=`),
                                   modification = trait.extinction(x = 0, condition = `>=`))
```

We can then feed these two extinction events to our previous simulations

```{r}
## Simulate the tree and traits with a random extinction event
sim_rand_extinction <- treats(
                   traits     = my_traits,
                   bd.params  = est_params,
                   stop.rule  = stop_rule,
                   events     = random_extinction,
                   null.error = 100,
                   replicates = 50)

## Simulate the tree and traits with a random extinction event
sim_trait_extinction <- treats(
                   traits     = my_traits,
                   bd.params  = est_params,
                   stop.rule  = stop_rule,
                   events     = negative_extinction,
                   null.error = 100,
                   replicates = 50)

## Visualising the difference between both scenarios
par(mfrow = c(1,2))
plot(sim_rand_extinction[[1]], main = "Random extinction")
plot(sim_trait_extinction[[1]], main = "Negative extinction")
```

We can then measure the disparity from these scenarios and compare them to our observed ones

```{r}
## Converting the data to be passed to dispRity functions
rand_ext_data <- dispRitreats(sim_rand_extinction)
trait_ext_data <- dispRitreats(sim_trait_extinction)

## Creating the time series
rand_ext_time_series <- chrono.subsets(data = rand_ext_data$data,
                                       tree = rand_ext_data$tree,
                                       time = seq(from = 0.9, to = 0, by = -0.1),
                                       method = "continuous", model = "proximity")
trait_ext_time_series <- chrono.subsets(data = trait_ext_data$data,
                                        tree = trait_ext_data$tree,
                                        time = seq(from = 0.9, to = 0, by = -0.1),
                                        method = "continuous", model = "proximity")

## Calculating the disparity from the simulated data
sim_rand_ext_disp <- dispRity(rand_ext_time_series, metric = c(sum, variances),
                              verbose = TRUE)
sim_trait_ext_disp <- dispRity(trait_ext_time_series, metric = c(sum, variances),
                               verbose = TRUE)

par(mfrow = c(2,2))
plot(change_in_size, main = "observed disparity")
plot(simulated_disparity, main = "simulated disparity")
plot(sim_rand_ext_disp, main = "simulated disparity\n(mass extinction)")
plot(simulated_disparity, main = "simulated disparity\n(trait extinction)")
```

Interestingly, with all our parameters used here and our specific disparity metric we cannot detect the effect of a mass extinction.
